---
title: "Chapter 6"
output: pdf_document
---

```{r echo=FALSE}
library(wooldridge)

```


## Ex.1 (wage1)

(i) Use OLS to estimate the following equation
\begin{equation}
  log(wage) = \beta_0 + \beta_1educ + \beta_2exper + \beta_3exper^2 + u 
\end{equation}
and report the results using the usual format.
(ii)  Is $exper^2$ statistically significant at the $1\%$ level?
(iii)  Using the approximation 
$$
  \%\Delta \hat{wage} \approx 100(\hat{\beta_2} +2\hat{\beta_3}exper)\Delta exper,
$$
     find the approximate return to the fifth year of experience? What is the approximate return to the twentieth year of experience?
(iv)  At what value of $exper$ does additional experience actually lower predicted log(wage)? How many people have more experience in this sample?

### Solution

(i)  The estimates of Equation 1 is
```{r,include=FALSE}
model1 <- lm(log(wage) ~ educ + exper + I(exper^2), data = wage1)
summary(model1)$coefficients
```

|        | Estimate | Std. Error | t value |
|------:|:-----|---------|:------:|
| Intercept | 0.1281 | 0.1059 | 1.2083 |
| $educ$ | 0.0937 | 0.0075 | 12.1004 |
| $exper$ | 0.0410 | 0.0052 | 7.8916 |
| $exper^2$ | -0.0007 | 0.0001 | -6.1639 |

(ii)  To check the statistical significance of $exper^2$, we compare the *t*-value of $exper^2$ with the critical *t*-value at $1\%$ significance level.  
  The $1\%$ critical value with 522 degrees of freedom is
  $$t^* = 2.576 < |t^*_{exper^2}| = 6.1639$$
 Therefore, $exper^2$ is statistically significant at 1\% level.  
 
(iii)  We have the following approximation
$$
  \%\Delta \hat{wage} \approx 100(\hat{\beta_2} +2\hat{\beta_3}exper)\Delta exper,
$$
  Plugging in $exper = 5$ in the above equation gives us the approximate return to 5 years of experience
\begin{equation}
\begin{split}
\frac{\%\Delta\hat{wage}}{\Delta exper} & = 100[0.0410+2(-0.007)(5)]\\
& \approx 3.387
\end{split}
\end{equation}
  This means that the wage increases by 3.387\% in going from 5 to 6 years of experience.  
  Similarly, the approximate return to the twentieth year of experience is
\begin{equation}
\begin{split}
\frac{\%\Delta\hat{wage}}{\Delta exper} & = 100[0.0410+2(-0.007)(20)]\\
& \approx 1.247
\end{split}
\end{equation}
 In going from twentieth years of experience to twenty-first year, the wage rate is expected to increase by 1.247\%.  
 
(iv)  The maximum years of experience that actually has a increasing effect on predicted wage rate, i.e., *turning point* is given by
\begin{equation}
\begin{split}
exper^* & = |\hat{\beta_2}/2\hat{\beta_3}|\\
& = |\frac{0.0410}{-0.0007}|\\
& = 5.85
\end{split}
\end{equation}
This means that after 6(5.85) years of experience, the return to log(wage) decreases. In the sample we have 390 people who have more than 6 years of experience.

## Ex.2 (wage2)

Consider a model where the return to education depends upon the amount of work experience (and vice versa):
\begin{equation}
log(wage) = \beta_0 + \beta_1educ + \beta_2exper + \beta_3educ.exper + u.
\end{equation}
(i)  Show that the return to another year of education (in decimal form), holding $exper$ fixed, is $\beta_1 + \beta_3exper$.  
(ii)  State the null hypothesis that the return to education does not depend on the level of $exper$. What do you think is the appropriate alternative?  
(iii)  Test the null hypothesis in (ii) against your stated alternative.  
(iv) Let $\theta_1$ denote the return to education (in decimal form), when $exper = 10:\theta_1 = \beta_1 + 10\beta_3$. Obtain the $\hat{\theta_1}$ and a 95\% confidence interval for $\theta_1$. (Hint: Write $\beta_1 = \theta_1 - 10\beta_3$ and plug this into the equation; then rearrange. This gives the regression for obtaining the confidence interval for $\theta_1$.)

### Solution

(i) 
\begin{equation}
\frac{\Delta{log(wage)}}{\Delta educ} = \beta_1educ + \beta_3exper
\end{equation}

